### 2012-03-26 exam4/2.tex
Exam 2H, problem 4.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{15}

Here are data and class definitions for representing the files and
directories of a computer:

\begin{verbatim}
;; An Elem is one of
;; - (new file% String Number)
;; - (new dir% String LoElem)
(define-class file% (fields name size))
(define-class dir% (fields name elems))

;; A LoElem is one of
;; - (new mt%)
;; - (new cons% Elem LoElem)
(define-class mt%)
(define-class cons% (fields first rest))    
\end{verbatim}

\noindent
Here is an example of a directory tree and its representation as an
\verb|Elem|:
\begin{verbatim}
#|
 User
   +-- A
   |   +-- a.txt
   |   +-- B
   |       +-- b.rkt
   +-- C
       +-- c.c
|#

(define mt (new mt%))
(define a (new file% "a.txt" 5))
(define b (new file% "b.rkt" 10))
(define c (new file% "c.c" 1000))

(define C/ (new dir% "C" (new cons% c mt)))
(define B/ (new dir% "B" (new cons% b mt)))
(define A/ (new dir% "A" (new cons% a (new cons% B/ mt))))
(define User/ (new dir% "/" (new cons% A/ (new cons% C/ mt))))
\end{verbatim}

\newpage
\noindent
Your job is to implement one of the main tasks of the common Unix
utility \verb|find|, which gives a list of all the files within an
element.  To do so, add a \verb|list-files| method of \verb|Elem|
which produces a list of files.  So for example, we expect:
\begin{verbatim}
(check-expect (User/ . list-files) 
              (new cons% a (new cons% b (new cons% c mt))))
\end{verbatim}
\emph{Hint}: it may come in handy to be able to concatenate two
\verb|LoElem| together.

\ifrubric
\else
\newpage
[Here is some more space for the previous problem.]
\fi

\begin{solution}
\begin{alltt}
\rubricpts{1pts -- interfaces, contracts, purpose}
;; An Elem implements
;; - list-files : -> LoElem
;;   List all files within this element.

\rubricpts{2pts -- interfaces, contracts, purpose}
;; A LoElem implements
;; - list-files : -> LoElem
;;   List all files within this list of elements.
;; - append : LoElem -> LoElem
;;   Append this list of elements to given list.

\rubricpts{1pt -- code}
;; in dir%
(define (list-files)
  (this . elems . list-files))

\rubricpts{1pt -- code}
;; in file%
(define (list-files)
  (new cons% this (new mt%)))

\rubricpts{2pt -- code}
;; in mt%
(define (append loe) loe)
(define (list-files) this)

\rubricpts{4pt -- code}
;; in cons%
(define (append loe)
  (new cons% (this . first) (this . rest . append loe)))
(define (list-files)
  (this . first . list-files . append (this . rest . list-files)))

\rubricpts{1pt -- sanity}

\rubricpts{3pts -- tests}
;; tests
(check-expect (mt . append (new cons% a mt)) (new cons% a mt))
(check-expect ((new cons% a mt) . append (new cons% b mt))
              (new cons% a (new cons% b mt)))

(check-expect (mt . list-files) mt)
(check-expect ((new cons% a (new cons% b mt)) . list-files)
              (new cons% a (new cons% b mt)))

(check-expect (a . list-files) (new cons% a mt))
\end{alltt}
\end{solution}

\end{problem}
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{18}
Last night, Asumu was up late improving the {\tt class} languages.
Unfortunately, he was too tired to test his work, and his last commit
broke the numeric system.
%
This puts us in the dire position of no longer being able to calculate
the factorial of large numbers, and we turn to you for help.

This is the definition of a class of functional objects
with a {\tt fact} method:

\begin{verbatim}
;; A Fact is (new fact%) and implements 
;; - fact : Natural -> Natural
;;   Compute n!.
(define-class fact%
  (define (fact n)
    (cond [(zero? n) 1]
          [else (* n (this . fact (sub1 n)))])))

(check-expect ((new fact%) . fact 0) 1)
(check-expect ((new fact%) . fact 5) 120)
\end{verbatim}

Since we can no longer rely on built in numbers to represent
\verb|Natural|s, we will represent \verb|Natural|s with objects and
then rewrite the \verb|fact| method to operate over this new kind of
data.

Here is the \verb|Natural| interface:

\begin{verbatim}
;; A Natural implements:
;; - zero? : -> Boolean
;;   is this Natural the zero value?
;; - sub1 : -> Natural
;;   produce a natural one less than this natural 
;;   (or zero if this natural is zero)
;; - times : Natural -> Natural
;;   multiply this natural by the given natural
\end{verbatim}

\newpage
Write a new definition of the \verb|fact| method using the
\verb|Natural| interface definition given above, and \emph{not} using
the built-in numbers of the class system or DrRacket.  You can assume
there is a defined constant \verb|one|, which is a \verb|Natural|
object representing $1$.

\begin{verbatim}
(check-expect ((new fact%) . fact one) one)
(check-expect ((new fact%) . fact (one . sub1)) one)
\end{verbatim}

\begin{solution}
\begin{alltt}
\rubricpts{2 pts}
(define-class fact%
  (define (fact n)
    (cond [(n . zero?) one]
          [else (n . times (this . fact (n . sub1)))])))

\end{alltt}
\end{solution}

Now, design data and class defintions that implement \verb|Natural|, and
supports the methods \verb|zero?|, \verb|sub1|, and \verb|times|.
Define the constant \verb|one|.  (\emph{Hint}: remember that the
natural numbers can be seen as a recursive union; a natural number is
either zero, or one more than a natural number.)

It will be helpful to define the helper methods \verb|add1| and
\verb|plus|.  Also, remember that $n + m = (n-1) + m + 1$ and that $n
\cdot m = (n-1) \cdot m + m$.  Don't overlook the slightly weird
specification for \verb|sub1|: $0-1 = 0$; this weirdness is common
amongst mathematicians when working with natural numbers.

\ifrubric
\else
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi
\end{problem}

\begin{solution}
\begin{alltt}
\rubricpts{2 pts -- data defintion}
;; A Natural is one of
;; - (new zero%)
;; - (new add1% Natural)
(define-class zero%
  \rubricpts{4 pts -- methods in zero%}
  (define (sub1) this)
  (define (zero?) true)
  (define (add1) (new add1% this))
  (define (plus n) n)
  (define (times n) this))

(define-class add1%
  (fields sub1)
  \rubricpts{7 pts -- methods in add1%}
  (define (zero?) false)
  (define (add1) (new add1% this))
  (define (plus n) (this . sub1 . plus n . add1))
  (define (times n) (this . sub1 . times n . plus n)))

(define zero (new zero%))
\rubricpts{1 pt -- define one}
(define one (new add1% (new zero%)))

\rubricpts{3 pts -- tests}
(check-expect (one . zero?) false)
(check-expect (one . sub1) zero)
(check-expect ((new add1% one) . sub1) one)
(check-expect (zero . zero?) true)
(check-expect (zero . sub1) zero)
\end{alltt}
\newpage
\end{solution}

\begin{problem}\pts{5}

\textbf{Extra Credit}
\\
\noindent
After Van Horn discovered the problems with numbers, he attempted to
fix them.  Unfortunately, not only did he not fix them, he broke
booleans as well.  Fortunately, we have an idea about how you can
help, expressed in the following interface definitions.

\begin{verbatim}
;; A Bool implements:
;; - branch : [Action X] [Action X] -> X
;;   run the first action if this bool is true, 
;;   otherwise run the second action

;; An [Action X] implements:
;; - run : -> X
;; run this action

\end{verbatim}

Following these interface definitions, design data and class
defintions for implementations of the \verb|Bool| interface, and
reimplement the \verb|fact| method to use them.
You will need to provide new defintions for \verb|true| and
\verb|false|, since they were broken by this catastrophe as well.

\ifrubric\else
\newpage
[Here is some more space for the previous problem.]
\fi

\begin{solution}
\begin{alltt}
\rubricpts{5 pts -- get it right}
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% one) (new recur% n))))

(define-class const%
  (fields run))

(define-class recur%
  (fields n)
  (define (run) 
    ((new fact%) . fact (this . n . sub1) . times (this . n))))

(define-class t%
  (define (branch t e) (t . run)))
(define-class f%
  (define (branch t e) (e . run)))

(define true (new t%)) 
(define false (new f%))

(check-expect ((new fact%) . fact zero) one)                  ; 0! = 1
(check-expect ((new fact%) . fact one) one)                   ; 1! = 1
(check-expect ((new fact%) . fact (one . add1)) (one . add1)) ; 2! = 2
(check-expect ((new fact%) . fact (one . add1 . add1))        ; 3! = 6
              (one . add1 . add1 . add1 . add1 . add1))
\end{alltt}
\end{solution}
\end{problem}
\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{??}

@@

\begin{problem} \pts{??}

Lists are nice, but sometimes it would be better to have a list
without an end.  For example, we might want to model the days of the
week as a list that wraps around at the end so that Monday follows
Sunday.  To do so, we need lists that are \emph{circular}, and in
order to make circular lists, we need to use mutation.  Here's an idea
for a list method that makes is possible to create cyclic lists:

\begin{verbatim}
interface List<X> {
   // EFFECT: set the rest of the last cons in this list
   // to the given list.
   void setTail(List<X> ls);
}
\end{verbatim}

\noindent
It works as follows:

\begin{verbatim}
List<Integer> waltz =
  new Cons<Integer>(1,
    new Cons<Integer>(2,
      new Cons<Integer>(3,
        new MT<Integer>())));

\end{verbatim}
At this point, {\tt waltz} is just a list of three elements {\tt 1, 2,
  3}.  To make the list circular we can call {\tt setTail} giving {\tt
  waltz} as the new tail:

\begin{verbatim}
waltz.setTail(waltz);
\end{verbatim}

\noindent
Now {\tt waltz} is a list of elements: {\tt 1, 2, 3, 1, 2, 3, 1, 2, 3,
  \dots}, and so on, \emph{ad infinitum}.

Of course, it's also possible to use {\tt setTail} to update a list in
a non-circular way.  Imagine instead we had just done:

\begin{verbatim}
waltz.setTail(new Cons<Integer>(4, new MT<Integer>()));
\end{verbatim}

\noindent
In this case, {\tt waltz} is now just {\tt 1, 2, 3, 4}.

\newpage
Design an implementation of {\tt setTail} that works as described.
Revise the {\tt List<X>} interface with any methods you need to add to
make {\tt setTail} work.  (You'll notice that the effect statement for
{\tt setTail} implies that this list cannot be empty; if you try to
{\tt setTail} on an empty list, you may raise an exception.)

\end{document}
